---
title: "Sample Size Justification"
runtime: shiny
output: 
    html_document:
        code_folding: hide
---

```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
library(magrittr)
library(pwr)
library(shiny)
knitr::opts_chunk$set(echo = FALSE)

rv <- reactiveValues(v_means = numeric())

getmode <- function(v) {
    uniqv <- unique(v)
    uniqv[which.max(tabulate(match(v, uniqv)))]
}
```

<style>
  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
</style>

# {.tabset}

## Variation in Single Samples

```{r v_variables}
fluidRow(
    column(2, numericInput("v_n", "Sample size (n)", value=10, min=1, step=1)),
    column(2, numericInput("v_mu", "Population mean", value=100)),
    column(2, numericInput("v_sd", "Population SD", value=15, min=0))
)

actionButton("v_button", label="Run simulation", icon("refresh"))
```

```{r v_run_simulation}
run_v_simulation <- eventReactive(input$v_button, {
    #create sample from normal distribution
    x <- rnorm(n = input$v_n, mean = input$v_mu, sd = input$v_sd)
    round_x <- round(x)
    # Store simulation mean
    rv$v_means <- c(rv$v_means, mean(x))

    # Create hypothetical normal distrubition using observed parameters
    sequence <- seq(min(x)/1.1, max(x)*1.1, length.out = 100)
    norm_vals <- dnorm(
        sequence, mean=mean(x), sd=sd(x)
    )
    # Normalise values so they plot nicely
    norm_vals <- data.table(
        x = sequence,
        y = norm_vals * length(round_x[round_x == getmode(round_x)]) /
          max(norm_vals)
    )

    list(
        x = x, sample_mean = mean(x), sample_sd = sd(x), mu=input$v_mu,
        sd = input$v_sd, n = input$v_n, norm_vals = norm_vals
    )
})
```

<div class="col2">

```{r v_plot}
renderText({if(input$v_button <= 0){print("Awaiting simulation...")}})
renderPlot({
    data <- run_v_simulation()
    ggplot() +
      geom_histogram(
          data=data.table(x=data$x), aes(x=x), binwidth=1, col="white"
      ) +
      geom_line(
          data=data$norm_vals, aes(x=x, y=y), col="red", linetype="dashed"
      ) +
      theme_bw() +
      labs(
          x="IQ",
          y="Count",
          title=paste0(
              data$n, " IQ samples from population with mean: ", data$mu,
              ", SD: ", data$sd
          ),
          subtitle=paste0(
              "Observed mean: ", round(data$sample_mean, 2),
              "\nObserved SD: ", round(data$sample_sd, 2)
          )
      )
})
```

```{r v_text}
output$v_text <- renderText({
    if(input$v_button > 0){
        paste0(
            "<b>Observed simulation means (ordered by size):</b><br>",
            paste0(round(rv$v_means[order(rv$v_means)], 2), collapse=", ")
        )
    }
})

htmlOutput("v_text")
```

</div>

## Sample Size Calculator

This calculates the number of participants required for each group for an
independent t-test, with specified alpha and required power level, to observe
a desired effect size.

```{r s_variables}
fluidRow(
    column(3, numericInput(
        "s_effect_size", "Expected effect side (d)", value=0.2, min=0, step=0.1)
    ),
    column(2, numericInput(
        "s_alpha", "Alpha level", value=0.05, min=0, step=0.01)
    ),
    column(2, numericInput(
        "s_power", "Required power", value=0.95, min=0, step=0.01)
    )
)
```

```{r s_samples required}
renderText({
    print(
        paste0(
            "Total samples required per group: ",
            round(
                pwr.t.test(
                    d=input$s_effect_size,
                    sig.level=input$s_alpha,
                    power=input$s_power
                )$n
            )
        )
    )
})
```


#

---

Daniel Lakens, 2016. 
This work is licensed under a Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International License. https://creativecommons.org/licenses/by-nc-sa/4.0/